\begin{thebibliography}{00}

\bibitem{ericsson-5g} Ericsson Mobility Report. [Online]. Available: \url{https://www.ericsson.com/en/mobility-report}. Accessed on Apr. 15, 2021.

\bibitem{iot-manifacturing} Chen, B., Wan, J., Celesti, A., Li, D., Abbas, H. and Zhang, Q., 2018. Edge computing in IoT-based manufacturing. IEEE Communications Magazine, 56(9), pp.103-109.

\bibitem{dsu} Mugarza, I., Amurrio, A., Azketa, E. and Jacob, E., 2019. Dynamic software updates to enhance security and privacy in high availability energy management applications in smart cities. IEEE Access, 7, pp.42269-42279.

\bibitem{saaas} Mitton, N., Papavassiliou, S., Puliafito, A. and Trivedi, K.S., 2012. Combining Cloud and sensors in a smart city environment.

\bibitem{iot-definition} Whitmore, A., Agarwal, A. and Da Xu, L., 2015. The Internet of Things—A survey of topics and trends. Information systems frontiers, 17(2), pp.261-274.

\bibitem{harnessing-continuum} Beckman, P., Dongarra, J., Ferrier, N., Fox, G., Moore, T., Reed, D. and Beck, M., 2020. Harnessing the Computing Continuum for Programming Our World. Fog Computing: Theory and Practice, pp.215-230.

\bibitem{computing-in-continuum} AbdelBaky, M., Zou, M., Zamani, A.R., Renart, E., Diaz-Montes, J. and Parashar, M., 2017, June. Computing in the continuum: Combining pervasive devices and services to support data-driven applications. In 2017 IEEE 37th International Conference on Distributed Computing Systems (ICDCS) (pp. 1815-1824). IEEE.

\bibitem{iot-enterprise} Haller S., Karnouskos S., Schroth C. (2009) The Internet of Things in an Enterprise Context. In: Domingue J., Fensel D., Traverso P. (eds) Future Internet – FIS 2008. FIS 2008. Lecture Notes in Computer Science, vol 5468. Springer, Berlin, Heidelberg.

\bibitem{fluid-internet} Latre, S., Famaey, J., De Turck, F. and Demeester, P., 2014. The fluid internet: service-centric management of a virtualised future internet. IEEE Communications Magazine, 52(1), pp.140-148.

\bibitem{cloud-iot} Botta, A., De Donato, W., Persico, V. and Pescapé, A., 2016. Integration of cloud computing and internet of things: a survey. Future generation computer systems, 56, pp.684-700.

\bibitem{cloud-def} Mell, P., and Grance, T. 2011. The NIST definition of cloud computing, Recommendations of the National Istitute of Standards and Technology, NIST Special Publication 800-145.

\bibitem{cloud-computing} Zhang, Q., Cheng, L. \& Boutaba, R. Cloud computing: state-of-the-art and research challenges. J Internet Serv Appl 1, 7–18 (2010).

\bibitem{gartner-iot} Gartner. Leading the IoT. 2017. [Online]. Available: \url{https://www.gartner.com/imagesrv/books/iot/iotEbook_digital.pdf}. Accessed on Apr. 15, 2021.

\bibitem{akamai} Nygren, E., Sitaraman, R.K. and Sun, J., 2010. The akamai network: a platform for high-performance internet applications. ACM SIGOPS Operating Systems Review, 44(3), pp.2-19.

\bibitem{aws-outage} Prolonged AWS outage takes down a big chunk of the internet. 2020. [Online]. Available: \url{https://www.theverge.com/2020/11/25/21719396/amazon-web-services-aws-outage-down-internet}. Accessed on: Apr. 15, 2021.

\bibitem{edge-computing-vision-challenges} Shi, W., Cao, J., Zhang, Q., Li, Y. and Xu, L., 2016. Edge computing: Vision and challenges. IEEE internet of things journal, 3(5), pp.637-646.

\bibitem{mobile-augmented-reality} Baresi, L., Mendonça, D.F. and Garriga, M., 2017, September. Empowering low-latency applications through a serverless edge computing architecture. In European Conference on Service-Oriented and Cloud Computing (pp. 196-210). Springer, Cham.

\bibitem{fog-computing} Cisco fog computing solutions: Unleash the power of the Internet of Things. [Online]. Available: \url{https://www.cisco.com/c/dam/en\_us/solutions/trends/iot/docs/computing-solutions.pdf}. Accessed on Apr. 15, 2021.

\bibitem{camera-virtualisation} Jang, S.Y., Lee, Y., Shin, B. and Lee, D., 2018, October. Application-aware IoT camera virtualisation for video analytics edge computing. In 2018 IEEE/ACM Symposium on Edge Computing (SEC) (pp. 132-144). IEEE.

\bibitem{raspberry} Raspberry Pi. [Online]. Available: \url{https://www.raspberrypi.org/products/}. Accessed on Apr. 15, 2021.

\bibitem{cloudlet} Satyanarayanan, M., Bahl, P., Caceres, R. and Davies, N., 2009. The case for vm-based cloudlets in mobile computing. IEEE pervasive Computing, 8(4), pp.14-23.

\bibitem{nfv-5g} Yousaf, F.Z., Bredel, M., Schaller, S. and Schneider, F., 2017. NFV and SDN—Key technology enablers for 5G networks. IEEE Journal on Selected Areas in Communications, 35(11), pp.2468-2478.

\bibitem{serverless-berkeley} Jonas, E., Schleier-Smith, J., Sreekanti, V., Tsai, C.C., Khandelwal, A., Pu, Q., Shankar, V., Carreira, J., Krauth, K., Yadwadkar, N. and Gonzalez, J.E., 2019. Cloud programming simplified: A berkeley view on serverless computing. arXiv preprint arXiv:1902.03383.

\bibitem{aws-lambda-edge} AWS Lambda@Edge. [Online]. Available: \url{https://aws.amazon.com/it/lambda/edge/}. Accessed on Apr. 15, 2021.

\bibitem{aws-greengrass} AWS IoT Greengrass. [Online]. Available: \url{https://aws.amazon.com/greengrass/}. Accessed on Apr. 15, 2021.

\bibitem{azure-iot-edge} Azure IoT Edge. [Online]. Available: \url{https://azure.microsoft.com/en-us/services/iot-edge/}. Accessed on Apr. 15, 2021.

\bibitem{serverless-mec} Baresi, L., Mendonça, D.F. and Garriga, M., 2017, September. Empowering low-latency applications through a serverless edge computing architecture. In European Conference on Service-Oriented and Cloud Computing (pp. 196-210). Springer, Cham.

\bibitem{oasis} MacKenzie, C.M., Laskey, K., McCabe, F., Brown, P.F., Metz, R. and Hamilton, B.A., 2006. Reference model for service oriented architecture 1.0. OASIS standard, 12(S 18).

\bibitem{protobuf} Protocol Buffers. [Online]. Available: \url{https://developers.google.com/protocol-buffers}. Accessed on Apr. 15, 2021.

\bibitem{openapi} OpenAPI Specification, Version 3.0.3. [Online]. Available: \url{https://swagger.io/specification/}. Accessed on Apr. 15, 2021.

\bibitem{microservices-today} Dragoni N. et al. (2017) Microservices: Yesterday, Today, and Tomorrow. In: Mazzara M., Meyer B. (eds) Present and Ulterior Software Engineering. Springer, Cham.

\bibitem{cloud-to-thing} T Lynn, JG Mooney, B Lee, PT Endo. 2020. The cloud-to-thing continuum: opportunities and challenges in cloud, fog and edge computing.

\bibitem{rest} Roy Fielding. 2000. Representational State Transfer (REST). \url{https://www.ics.uci.edu/~fielding/pubs/dissertation/rest\_arch\_style.htm}. Accessed on: Apr. 15, 2021.

\bibitem{shopify} Shopify App Store. [Online]. Available: \url{https://apps.shopify.com/}. Accessed on Apr. 15, 2021.

\bibitem{vmm} Rosenblum, M. and Garfinkel, T., 2005. Virtual machine monitors: Current technology and future trends. Computer, 38(5), pp.39-47.

\bibitem{kvm} Kernel Virtual Machine. [Online]. Available: \url{https://www.linux-kvm.org/page/Main_Page}. Accessed on Apr. 15, 2021.

\bibitem{containers} Bernstein, D., 2014. Containers and cloud: From lxc to docker to kubernetes. IEEE Cloud Computing, 1(3), pp.81-84.

\bibitem{containers-raspberry} Pahl, C., Helmer, S., Miori, L., Sanin, J. and Lee, B., 2016, August. A container-based edge cloud paas architecture based on raspberry pi clusters. In 2016 IEEE 4th International Conference on Future Internet of Things and Cloud Workshops (FiCloudW) (pp. 117-124). IEEE.

\bibitem{containers-fog} Bellavista, P. and Zanni, A., 2017, January. Feasibility of fog computing deployment based on docker containerization over raspberrypi. In Proceedings of the 18th international conference on distributed computing and networking (pp. 1-10).

\bibitem{gcloud-functions} Google Cloud Functions. [Online]. Available: \url{https://cloud.google.com/functions}. Accessed on Apr. 15, 2021.

\bibitem{azure-functions} Azure Functions. [Online]. Available: \url{https://azure.microsoft.com/en-us/services/functions/}. Accessed on Apr. 15, 2021.

\bibitem{aws-lambda} AWS Lambda. [Online]. Available: \url{https://aws.amazon.com/lambda/}. Accessed on Apr. 15, 2021.

\bibitem{serverless-evaluation} Mohanty, S.K., Premsankar, G. and Di Francesco, M., 2018, December. An Evaluation of Open Source Serverless Computing Frameworks. In CloudCom (pp. 115-120).

\bibitem{urllc} Elbamby, M.S., Perfecto, C., Liu, C.F., Park, J., Samarakoon, S., Chen, X. and Bennis, M., 2019. Wireless edge computing with latency and reliability guarantees. Proceedings of the IEEE, 107(8), pp.1717-1737.

\bibitem{gvisor-firecracker} Caraza-Harter, T. and Swift, M.M., 2020, March. Blending containers and virtual machines: a study of firecracker and gVisor. In Proceedings of the 16th ACM SIGPLAN/SIGOPS International Conference on Virtual Execution Environments (pp. 101-113).

\bibitem{firecracker} Firecracker. [Online]. Available: \url{https://firecracker-microvm.github.io/}. Accessed on Apr. 15, 2021.

\bibitem{unikernels} Madhavapeddy, A., Mortier, R., Rotsos, C., Scott, D., Singh, B., Gazagnaire, T., Smith, S., Hand, S. and Crowcroft, J., 2013. Unikernels: Library operating systems for the cloud. ACM SIGARCH Computer Architecture News, 41(1), pp.461-472.

\bibitem{unikernels-performance} Koller, R. and Williams, D., 2017, May. Will serverless end the dominance of linux in the cloud?. In Proceedings of the 16th Workshop on Hot Topics in Operating Systems (pp. 169-173).

\bibitem{sfi} Wahbe, R., Lucco, S., Anderson, T.E. and Graham, S.L., 1993, December. Efficient software-based fault isolation. In Proceedings of the fourteenth ACM symposium on Operating systems principles (pp. 203-216).

\bibitem{mqtt-coap-amqp-http} Naik, N., 2017, October. Choice of effective messaging protocols for IoT systems: MQTT, CoAP, AMQP and HTTP. In 2017 IEEE international systems engineering symposium (ISSE) (pp. 1-7). IEEE.

\bibitem{rest-opc} Grüner, S., Pfrommer, J. and Palm, F., 2016. RESTful industrial communication with OPC UA. IEEE Transactions on Industrial Informatics, 12(5), pp.1832-1841.

\bibitem{inter-cloud} Grozev, N. and Buyya, R., 2014. Inter‐Cloud architectures and application brokering: taxonomy and survey. Software: Practice and Experience, 44(3), pp.369-390.

\bibitem{docker-hub} Docker Hub. [Online]. Available: \url{https://hub.docker.com/}. Accessed on Apr. 15, 2021.

\bibitem{lavea} Yi, S., Hao, Z., Zhang, Q., Zhang, Q., Shi, W. and Li, Q., 2017, October. Lavea: Latency-aware video analytics on edge computing platform. In Proceedings of the Second ACM/IEEE Symposium on Edge Computing (pp. 1-13).

\bibitem{map-reduce} Condie, T., Conway, N., Alvaro, P., Hellerstein, J.M., Elmeleegy, K. and Sears, R., 2010, April. MapReduce online. In Nsdi (Vol. 10, No. 4, p. 20).

\bibitem{data-analytics-iot} He, J., Wei, J., Chen, K., Tang, Z., Zhou, Y. and Zhang, Y., 2017. Multitier fog computing with large-scale iot data analytics for smart cities. IEEE Internet of Things Journal, 5(2), pp.677-686.

\bibitem{vehicular-data-cloud} He, W., Yan, G. and Da Xu, L., 2014. Developing vehicular data cloud services in the IoT environment. IEEE transactions on industrial informatics, 10(2), pp.1587-1595.

\bibitem{edge-computing-survey} Yu, W., Liang, F., He, X., Hatcher, W.G., Lu, C., Lin, J. and Yang, X., 2017. A survey on the edge computing for the Internet of Things. IEEE access, 6, pp.6900-6919.

\bibitem{quic} Langley, A., Riddoch, A., Wilk, A., Vicente, A., Krasic, C., Zhang, D., Yang, F., Kouranov, F., Swett, I., Iyengar, J. and Bailey, J., 2017, August. The quic transport protocol: Design and internet-scale deployment. In Proceedings of the conference of the ACM special interest group on data communication (pp. 183-196).

\bibitem{coap} Bormann, C., Castellani, A.P. and Shelby, Z., 2012. Coap: An application protocol for billions of tiny internet nodes. IEEE Internet Computing, 16(2), pp.62-67.

\bibitem{web-of-things} Guinard, D., Trifa, V. and Wilde, E., 2010, November. A resource oriented architecture for the web of things. In 2010 Internet of Things (IOT) (pp. 1-8). IEEE.

\bibitem{cloud-foundry-marketplace} Cloud Foundry Launches Online Marketplace for Expanding Ecosystem. [Online]. Available: \url{https://www.cloudfoundry.org/blog/cloud-foundry-launches-online-marketplace-expanding-ecosystem/}. Accessed on Apr. 15, 2021.

\bibitem{osb} Open Service Broker. [Online]. Available: \url{https://www.openservicebrokerapi.org/}. Accessed on Apr. 15, 2021.

\bibitem{kubernetes} Kubernetes. [Online]. Available: \url{https://kubernetes.io/}. Accessed on Apr. 15, 2021.

\bibitem{containerd} containerd. [Online]. Available: \url{https://containerd.io/}. Accessed on Apr. 15, 2021.

\bibitem{docker-swarm} Docker Swarm overview. [Online]. Available: \url{https://docs.docker.com/engine/swarm/}. Accessed on Apr. 15, 2021.

\bibitem{akri} Akri. [Online]. Available: \url{https://github.com/deislabs/akri}. Accessed on Apr. 15, 2021.

\bibitem{k8s-device-plugins} Kubernetes Device Plugins. [Online]. Available: \url{https://kubernetes.io/docs/concepts/extend-kubernetes/compute-storage-net/device-plugins/}. Accessed on Apr. 15, 2021.

\bibitem{gRPC} gRPC. A high performance, open source universal RPC framework. [Online]. Available: \url{https://grpc.io/}. Accessed on Apr. 15, 2021.

\bibitem{wasm} Haas, A., Rossberg, A., Schuff, D.L., Titzer, B.L., Holman, M., Gohman, D., Wagner, L., Zakai, A. and Bastien, J.F., 2017, June. Bringing the web up to speed with WebAssembly. In Proceedings of the 38th ACM SIGPLAN Conference on Programming Language Design and Implementation (pp. 185-200).

\bibitem{typescript} TypeScript. [Online]. Available: \url{https://www.typescriptlang.org/}. Accessed on Apr. 15, 2021.

\bibitem{npm-security} Microsoft spots malicious npm package stealing data from UNIX systems. [Online]. Available: \url{https://www.zdnet.com/article/microsoft-spots-malicious-npm-package-stealing-data-from-unix-systems/}. Accessed on Apr. 15, 2021.

\bibitem{reference-types} Reference Types Overview. [Online]. Available: \url{https://github.com/WebAssembly/reference-types/blob/master/proposals/reference-types/Overview.md}. Accessed on Apr. 15, 2021.

\bibitem{llvm} LLVM. [Online]. Available: \url{https://llvm.org/}. Accessed on Apr. 15, 2021.

\bibitem{cloudflare-workers} WebAssembly on Cloudflare Workers. [Online]. Available: \url{https://blog.cloudflare.com/webassembly-on-cloudflare-workers/}. Accessed on Apr. 15, 2021.

\bibitem{fastly-lucet} Announcing Lucet: Fastly's native WebAssembly compiler and runtime. [Online]. Available: \url{https://www.fastly.com/blog/announcing-lucet-fastly-native-webassembly-compiler-runtime}. Accessed on Apr. 15, 2021.

\bibitem{parity-wasm} Write Wasm smart contracts with ink! 2.0. [Online]. Available: \url{https://www.parity.io/write-wasm-smart-contracts-with-ink-2-0/}. Accessed on Apr. 15, 2021.

\bibitem{sledge} Gadepalli, P.K., McBride, S., Peach, G., Cherkasova, L. and Parmer, G., 2020, December. Sledge: a Serverless-first, Light-weight Wasm Runtime for the Edge. In Proceedings of the 21st International Middleware Conference (pp. 265-279).

\bibitem{execution-model-serverless-edge} Hall, A. and Ramachandran, U., 2019, April. An execution model for serverless functions at the edge. In Proceedings of the International Conference on Internet of Things Design and Implementation (pp. 225-236).

\bibitem{faasm} Shillaker, S. and Pietzuch, P., 2020. Faasm: lightweight isolation for efficient stateful serverless computing. In 2020 {USENIX} Annual Technical Conference ({USENIX}{ATC} 20) (pp. 419-433).

\bibitem{wasm3} wasm3. [Online]. Available: \url{https://github.com/wasm3/wasm3}. Accessed on Apr. 15, 2021.

\bibitem{ewasm} Peach, G., Pan, R., Wu, Z., Parmer, G., Haster, C. and Cherkasova, L., 2020. eWASM: Practical Software Fault Isolation for Reliable Embedded Devices. IEEE Transactions on Computer-Aided Design of Integrated Circuits and Systems, 39(11), pp.3492-3505.

\bibitem{wasmtime} wasmtime. [Online]. Available: \url{https://github.com/bytecodealliance/wasmtime}. Accessed on Apr. 15, 2021.

\bibitem{wasm3-performance} wasm3 performance. [Online]. Available: \url{https://github.com/wasm3/wasm3/blob/main/docs/Performance.md}. Accessed on Apr. 15, 2021.

\bibitem{levee} Brzoza-Woch, R., Konieczny, M., Nawrocki, P., Szydlo, T. and Zielinski, K., 2016, May. Embedded systems in the application of fog computing—Levee monitoring use case. In 2016 11th IEEE Symposium on Industrial Embedded Systems (SIES) (pp. 1-6). IEEE.

\bibitem{coap-soa} Pereira, P.P., Eliasson, J., Kyusakov, R., Delsing, J., Raayatinezhad, A. and Johansson, M., 2013, March. Enabling cloud connectivity for mobile internet of things applications. In 2013 IEEE seventh international symposium on service-oriented system engineering (pp. 518-526). IEEE.

\bibitem{go-data-races} Golang data races to break memory safety. [Online]. Available: \url{https://blog.stalkr.net/2015/04/golang-data-races-to-break-memory-safety.html}. Accessed on Apr. 15, 2021.

\bibitem{stack-resource-policy} Baker, T.P., 1990, December. A stack-based resource allocation policy for realtime processes. In [1990] Proceedings 11th Real-Time Systems Symposium (pp. 191-200). IEEE.

\bibitem{definitive-guide-cortex} Joseph, Y., 2014. The definitive guide to ARM Cortex-M3 and Cortex-M4 processors.

\bibitem{interrupts-handling} Interrupt inputs and pending behaviors, Y., 2014. The definitive guide to ARM Cortex-M3 and Cortex-M4 processors. pp 246.

\bibitem{iot-healthcare} Pace, P., Aloi, G., Gravina, R., Caliciuri, G., Fortino, G. and Liotta, A., 2018. An edge-based architecture to support efficient applications for healthcare industry 4.0. IEEE Transactions on Industrial Informatics, 15(1), pp.481-489.

\bibitem{gartner-cloud-edge} Gartner. Predicts 2021: Cloud and Edge Infrastructure. [Online]. Available: \url{https://www.gartner.com/en/doc/735107-predicts-2021-cloud-and-edge-infrastructure}. Accessed on Apr. 15, 2021.

\bibitem{downtime-cost} Cost of Data Center Outages. January 2016. [Online]. Available: \url{http://files.server-rack-online.com/2016-Cost-of-Data-Center-Outages.pdf}. Accessed on Apr. 15, 2021.

\bibitem{traceroute-google} Visual Traceroute. [Online]. Available: \url{https://www.dotcom-monitor.com/wiki/knowledge-base/visual-traceroute-graphical-tool/}. Accessed on Apr. 15, 2021.

\bibitem{mec} Beck, M.T., Werner, M., Feld, S. and Schimper, S., 2014, November. Mobile edge computing: A taxonomy. In Proc. of the Sixth International Conference on Advances in Future Internet (pp. 48-55). Citeseer.

\bibitem{docker-image} Docker Image Specification 1.0.0. [Online]. Available: \url{https://github.com/moby/moby/blob/master/image/spec/v1.md}. Accessed on Apr. 15, 2021.

\bibitem{gcloud-function-build} Building Cloud Functions Images. [Online]. Available: \url{https://cloud.google.com/functions/docs/building}. Accessed on Apr. 15, 2021.

\bibitem{rust} Rust. [Online]. Available: \url{https://www.rust-lang.org/}. Accessed on Apr. 15, 2021.

\bibitem{k8s-architecture} Introduction to Kubernetes architecture. [Online]. Available: \url{https://www.redhat.com/en/topics/containers/kubernetes-architecture}. Accessed on Apr. 15, 2021.

\bibitem{etcd} etcd. [Online]. Available: \url{https://etcd.io/}. Accessed on Apr. 15, 2021.

\bibitem{docker-edge-computing} Ismail, B.I., Goortani, E.M., Ab Karim, M.B., Tat, W.M., Setapa, S., Luke, J.Y. and Hoe, O.H., 2015, August. Evaluation of docker as edge computing platform. In 2015 IEEE Conference on Open Systems (ICOS) (pp. 130-135). IEEE.
\bibitem{etcd} etcd. [Online]. Available: \url{https://etcd.io/}. Accessed on Apr. 15, 2021.

\bibitem{onvif} ONVIF. [Online]. Available: \url{https://www.onvif.org/}. Accessed on Apr. 15, 2021.

\bibitem{udev} udev. [Online]. Available: \url{https://wiki.archlinux.org/index.php/udev}. Accessed on Apr. 15, 2021.

\bibitem{rtic} RTIC. [Online]. Available: \url{https://rtic.rs/0.5/}. Accessed on Apr. 15, 2021.

\bibitem{rtfm} Eriksson, J., Häggström, F., Aittamaa, S., Kruglyak, A. and Lindgren, P., 2013, June. Real-time for the masses, step 1: Programming API and static priority SRP kernel primitives. In 2013 8th ieee international symposium on industrial embedded systems (sies) (pp. 110-113). IEEE.

\end{thebibliography}
